---
title: "Aid Success"
author: "Anwesha Guha"
date: "5/27/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup & pull in data

```{r}
library(pacman)
p_load(tidyverse, rio, here, janitor, psych)
```

Read in individual files.
```{r}
act_sat_conversion <- import(here("data", "ACT_SAT_concordance.csv")) %>% 
  clean_names()

hs_gpa_test <- import(here("data", "clean_hs_gpa_different_test_score_4ag.csv")) %>% 
  clean_names() #19964 NA values for converted
first_enroll <- import(here("data", "first_term_enrolled_4ag.csv"))  %>% 
  clean_names()#46496 NA values for ID, only ID & first term enrolled

gpa <- import(here("data", "gpa_4ag.csv")) %>% 
  clean_names() #college gpa info

hs <- import(here("data", "highschool_4ag.csv")) %>% 
  clean_names()

med_income <- import(here("data", "median_household_income_4ag.csv"))  %>% 
  clean_names()#smallest N

sat_old_new <- import(here("data", "SAT_old_new_concordance.csv")) %>% 
  clean_names()

sex_race_path <- import(here("data", "sex_racethn_pway_4ag.csv"))  %>% 
  clean_names()#not as updated, so ignore

sex_race_res_path <- import(here("data", "sex_racethn_rsdncy_pway_4ag.csv")) %>% 
  clean_names()

swasi_demo <- import(here("data", "swasi_bl_wb_and_demos_4ag.csv")) %>% 
  clean_names()

test <- import(here("data", "test_4ag.csv")) %>% 
  clean_names()
```

Combine following data into one file.

* sex_race_res_path (id, academic_period_admitted)
* gpa (id, academic_period_admitted)
* hs (id, academic_period_admitted)

```{r}
success1 <- left_join(gpa, sex_race_res_path)
```

Since we want to filter on first-time college entrants, we filter students who have legitimate HS information only. 
```{r}
hs_only <- hs %>% 
  filter(institution_type == "H") %>% 
  select(-c(institution_type, transfer_credits_attendance))

hs_only$school_gpa <- str_replace_all(hs_only$school_gpa, ",", ".")
hs_only$school_gpa <- as.numeric(hs_only$school_gpa)

hs_only <- hs_only %>% 
  filter(school_gpa >= 1 & school_gpa < 5) %>% 
  mutate(id_period = paste0(id, academic_period_admitted)) 

# keep highest GPA value when there are multiple HS reported 
# this is to avoid a many-to-many join
hs_only <- hs_only[order(hs_only$id_period, -hs_only$school_gpa), ]
hs_only <- hs_only[ !duplicated(hs_only$id_period), ]
```

```{r}
success <- left_join(hs_only, success1) %>% 
  select(-id_period)

# write_csv(success, "success.csv")
```

## Descriptives

```{r}
describe(success)
```

```{r}
hist(success$school_gpa)
```


```{r}
success <- success %>% 
  filter(gpa>0)

hist(success$gpa)
```


```{r}
table(success$gender)
```

```{r}
table(success$federal_ethnic_desc)
```

```{r}
table(success$residency_desc)

success_res <- success %>% 
  filter(residency_desc == "Resident") %>% 
  select(-residency_desc) 
```

This brings us down to `r length(unique(success_res$id))` unique IDs. 

Let's look at descriptives of Pathway scholars.

```{r}
pathway_only <- success_res %>% 
  filter(student_attribute == "PW")
```

```{r}
hist(pathway_only$school_gpa)
```

```{r}
hist(pathway_only$gpa)
```


```{r}
table(pathway_only$gender)
```

```{r}
table(pathway_only$federal_ethnic_desc)
```

Let's look at these side-by-side.

```{r}
success_pathway_viz <- success_res %>% 
  mutate(pathway_status = case_when(student_attribute == "PW" ~ "Pathway",
                                    TRUE ~ "No Pathway"))
```


```{r}
boxplot(success_pathway_viz$school_gpa ~ success_pathway_viz$pathway_status,
        horizontal = TRUE)
```

```{r}
boxplot(success_pathway_viz$gpa ~ success_pathway_viz$pathway_status,
        horizontal = TRUE)
```

I'd really like to look at family/household income descriptors also.

